# Data transformation

Author: yueyanw  
Reviewer: Chunji

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(nycflights13)
```
### 5.2.4 Exercises
1. Find all flights that
    1. Had an arrival delay of two or more hours
        ```{r}
        filter(flights, arr_delay >= 120)
        ```
    2. Flew to Houston (`IAH` or `HOU`)
        ```{r}
        filter(flights, dest %in% c('IAH', 'HOU'))
        ```
    3. Were operated by United, American, or Delta
    4. Departed in summer (July, August, and September)
    5. Arrived more than two hours late, but didn’t leave late
    6. Were delayed by at least an hour, but made up over 30 minutes in flight
    7. Departed between midnight and 6am (inclusive)



5.3.1.1 How could you use `arrange()` to sort all missing values to the start? (Hint: use `is.na()`).

```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, !is.na(x), x)
```

5.6.7.2

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% count(tailnum, wt = distance)
not_cancelled %>% group_by(tailnum) %>% summarize(n = sum(distance))
```
