# Data transformation

Author: yueyanw  
Reviewer: Chunji

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(nycflights13)
```
### 5.2.4 Exercises
1. Find all flights that
    1. Had an arrival delay of two or more hours
        ```{r}
        filter(flights, arr_delay >= 120)
        ```
    2. Flew to Houston (`IAH` or `HOU`)
        ```{r}
        filter(flights, dest %in% c('IAH', 'HOU'))
        ```
    3. Were operated by United, American, or Delta
    We can look up the 2-character `carrier` code in the `airlines` dataset.
        ```{r}
        airlines
        ```
        United is `UA`, American is `AA`, and Delta is `DL`.
        ```{r}
        filter(flights, carrier %in% c('UA','AA','DL'))
        ```
    4. Departed in summer (July, August, and September)
        ```{r}
        filter(flights, month %in% c(7,8,9))
        ```
    5. Arrived more than two hours late, but didn't leave late
        ```{r}
        filter(flights, arr_delay > 120 & dep_delay == 0)
        ```
    6. Were delayed by at least an hour, but made up over 30 minutes in flight
        ```{r}
        filter(flights, dep_delay >= 60 & dep_delay - arr_delay >=30)
        ```
    7. Departed between midnight and 6am (inclusive)
    Midnight is recorded as `2400` instead of `0`
        ```{r}
        filter(flights, dep_time == 2400 | (dep_time >= 0000 & dep_time <= 0600))
        ```
2. Another useful dplyr filtering helper is `between()`. What does it do? Can you use it to simplify the code needed to answer the previous challenges?
`between()` is a shortcut for `x >= left & x <= right`. Challenges 4. and 7. can be simplified as follows:
    ```{r}
    filter(flights, between(month, 7,9))
    filter(flights, dep_time == 2400|between(dep_time, 0000,0600))
    ```
3. How many flights have a missing `dep_time`? What other variables are missing? What might these rows represent?
8255 flights have a missing `dep_time`. `dep_delay`,`arr_time`, `arr_delay`, and `air_time` are also missing. These rows might represent cancelled flights.
    ```{r}
    sum(is.na(flights$dep_time))
    summary(filter(flights, is.na(dep_time) == TRUE))
    ```
4. Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing? Why is `FALSE & NA` not missing? Can you figure out the general rule? (`NA * 0` is a tricky counterexample!)
`NA ^ 0` is not missing, because any number to the zero power is 1.
    ```{r}
    NA ^ 0
    ```
`NA | TRUE` is not missing, because `TRUE` or any logical value always returns as `TRUE`. 
    ```{r}
    NA | TRUE
    ```
`FALSE & NA` is not missing, because `FALSE` and any logical always returns as `FALSE`. 
    ```{r}
    FALSE & NA
    ```
The general rule is that `NA` is treated as a placeholder for a logical value or numeric value. If the operation returns to a value regardless any possible value `NA` might take, it will return that value. 
`NA * 0` returns `NA`, which seems counterintuitive, because we know that any number times 0 should return a 0. 
    ```{r}
    NA * 0
    ```
However, there is an exception. x * 0 is undefined when `x == Inf` or `x == -Inf`.
    ```{r}
    Inf * 0
    -Inf *0
    ```

5.3.1.1 How could you use `arrange()` to sort all missing values to the start? (Hint: use `is.na()`).

```{r}
df <- tibble(x = c(5, 2, NA))
arrange(df, !is.na(x), x)
```

5.6.7.2

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% count(tailnum, wt = distance)
not_cancelled %>% group_by(tailnum) %>% summarize(n = sum(distance))
```
